---
title: "Moring Consult Part B"
author: "John Warren "
date: '2022-05-06'
output: md_document
---

```{r}
rm(list=ls()) #rm(list of objects) removes all objects from memory
graphics.off() #Closing all previously open graphs
```

```{r,message = FALSE, warning = FALSE }
library(tidyverse)
library(lubridate)
library(dplyr)
library("data.table")
library(readr)
library(car)
library(readxl)
library(mice)
library(gapminder)
library(here)
library(ggplot2)
#library(ggplot)
```


```{r }
mtcars <- read_csv("C:/Users/Admin/Downloads/mtcars.csv")
```


```{r}

ggplot(mtcars, aes(x=make, y=mpg)) + 
  geom_bar(aes(reorder(make,mpg),mpg),stat="identity", width=.5, fill="Red") + 
  labs(title="Make Vs mpg") + 
  theme(axis.text.x = element_text(angle=70, vjust=0.6))

```

The graphic above shows cars make and their mpg. The top 5 cars that get the best MPG are Toyota Corolla, Fiat 128, lotus Europa, Honda civic, and the FiatX1_9. The cars with the lowest MPG in the data set are Chrysler Imperial, Duster 360, Camaro Z28, and Lincoln Continental and the last with the worst MPG is the Cadillac Fleetwood.

```{r}
#importing packages

library(ggcorrplot)
library(ggthemes)

# Correlation matrix
data(mtcars)
corr <- round(cor(mtcars), 1)

# Ploting Corelation matrix of mtcars
ggcorrplot(corr, hc.order = TRUE, 
           type = "lower", 
           lab = TRUE, 
           lab_size = 3, 
           colors = c("red", "white", "green"), 
           title="Corelation matrix of mtcars", 
           ggtheme=theme_bw)
```

This heat map represent the relationship between different features of the car.For example we can see that a cars weight is negatively correlated with MPG at -.9 this can be seen in the graphic on the bottom right hand Conner. 


lets run a regression

In this model, I use an OLS analysis technique to understand the relationship between a car's weight and the MPG that it gets on average when driving on the freeway. My null hypothesis is that a car's weight will negatively affect its mpg.
```{r}
model<- lm(mpg~wt,data = mtcars)
summary(model)
plot(wt~mpg, data=mtcars, main="Wight vs Mpg", col="blue",lwd=2)
legend("topright", legend = c("Cars"),
       col=c("blue","red"), lty=c(0,1), pch=c(1,NA),lwd=2)

```

In the regression results, we can see with a car’s weight of 0 the mpg is stated to be 37 miles per gallon the intercept rarely makes any sense, therefore, we will leave it at that and use it as a starting point. When we analyze the coefficient estimates results for the weight of a car, we can see that it has a negative relationship with miles per gallon the P-value on this coefficient reaches the highest level of significance which allows us to confirm that this variable does have significance within our model. when we look at the coefficient estimates results, we can see that with every one unit increase in a car weight we should see a fall in a cars MPG by 5.34 miles. When looking at the r ^2 of this model we can see that it encompasses about 75% of the variation inside this data set. In this is a simple regression model we can assume that there might be omitted variable bias hiding inside of the error term as we include more variables into our model will be able to paint a better picture of what contributes to a car’s miles per gallon.

Cleaning the environment for new Data 
```{r}
rm(list=ls()) #rm(list of objects) removes all objects from memory
graphics.off() #Closing all previously open graphs
```

Part 2 data ViZ
```{r}
viz <- read_csv("C:/Users/Admin/OneDrive/Desktop/morning conslut/part 2 data.csv")
```


Cleaning up the data 
```{r}
viz <- viz %>% na.omit()

```

```{r}
View(viz)
sapply(viz,class)

```
Recreating the data Viz 
```{r}
a<- viz %>% pivot_longer(-Demographic) 
a %>% group_by(Demographic)%>% mutate(pvalue = round(value/sum(value),2)) %>% ungroup() %>%
ggplot(aes(y = pvalue, x = name, fill = name))+ geom_col() + facet_wrap(~Demographic) +
scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())+
  labs( x = "Opinion", y = "Percept", fill = "Opinion")

  
```


